---
title: "How to Datapasta"
author: "Miles McBain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Datapasta}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Datapasta provides three Rstudio addins that give you complete freedom copy and paste data into your RStudio scource editor, formatted for immeddiate use. Note: repeated use has been known to cause **titilation** and **giddiness**.

Places I've found this power useful:

* Copying tables from Excel, Jupyter, and websites, where the source file cannot be easily read.
* Embedding small-ish amounts of raw data from .csv into Rmarkdown files. The file thus constains code documentation and data, attaining the holy trinity of reproducibility.
* Quickly pasting vector output from other queries into `dplyr::filter( .. %in% ..)`.
* Adding datasets to readily reproducible examples for posting to StackOverflow, Slack channels etc.

The addins and their behaviour are now described.

## Pasting a table as a formatted tibble definition with `tribble_paste()`
You can copy this html table of Brisbane weather forecasts:

```{r, echo=FALSE}
library(tibble)
knitr::kable(tribble(
                                  ~X,          ~Location, ~Min, ~Max,
                    "Partly cloudy.",         "Brisbane",   19,   29,
                    "Partly cloudy.", "Brisbane Airport",   18,   27,
                  "Possible shower.",       "Beaudesert",   15,   30,
                    "Partly cloudy.",        "Chermside",   17,   29,
    "Shower or two. Possible storm.",           "Gatton",   15,   32,
                  "Possible shower.",          "Ipswich",   15,   30,
                    "Partly cloudy.",    "Logan Central",   18,   29,
                     "Mostly sunny.",            "Manly",   20,   26,
                    "Partly cloudy.",    "Mount Gravatt",   17,   28,
                  "Possible shower.",            "Oxley",   17,   30,
                    "Partly cloudy.",        "Redcliffe",   19,   27
    ))
```

And make this appear in your source editor, at the current cursor:

```{r eval=FALSE}
tribble(
                                  ~X,          ~Location, ~Min, ~Max,
                    "Partly cloudy.",         "Brisbane",   19,   29,
                    "Partly cloudy.", "Brisbane Airport",   18,   27,
                  "Possible shower.",       "Beaudesert",   15,   30,
                    "Partly cloudy.",        "Chermside",   17,   29,
    "Shower or two. Possible storm.",           "Gatton",   15,   32,
                  "Possible shower.",          "Ipswich",   15,   30,
                    "Partly cloudy.",    "Logan Central",   18,   29,
                     "Mostly sunny.",            "Manly",   20,   26,
                    "Partly cloudy.",    "Mount Gravatt",   17,   28,
                  "Possible shower.",            "Oxley",   17,   30,
                    "Partly cloudy.",        "Redcliffe",   19,   27
    )
```

`tibble::tribble()` or '**transposed** tibble' is a really neat function that allows a `tibble` to be written in human readable format (Thanks be to Hadley).

To paste data as a `tribble()` call, just copy the table header and data rows, then paste into the source editor using the addin `Paste as tribble`. For best results, assign the addin to a memorable keyboard shortcut, e.g. `ctrl + shift + t`. See [Customizing Keyboard Shortcuts ](https://support.rstudio.com/hc/en-us/articles/206382178-Customizing-Keyboard-Shortcuts).

`tribble_paste()` is a flexible function that guesses the separator and types of the data it pulls from the clipboard. Mostly this seems to work well. Ocassionally it epic-fails. The supported separators are `\|` (pipe), `\t` (tab), `,` (comma), `;`(semicolon). Most data copied from the internet or spreadsheets will be tab delimited.

##Pasting a list as a horizontal vector with `vector_paste()`
A list could be a row or column of a spreadsheet or intermediate output. With the `Paste as vector` addin you can go from something like:

```
Mint	Fedora	Debian	Ubuntu	OpenSUSE
```
or
```
Mint,Fedora,Debian,Ubuntu,OpenSUSE
```
or
```
Mint
Fedora
Debian
Ubuntu
OpenSUSE
```
to

```{r, eval = FALSE}
c("Mint", "Fedora", "Debian", "Ubuntu", "OpenSUSE")
```

This is pasted into the souce editor at the current cursor.

Just like `tribble_paste()`, `vector_paste()` has a flexible parser that can guess the type and separator of the data. The supported separators are `\|` (pipe), `\t` (tab), `,` (comma), `;`(semicolon) and end of line. The recommended keyboard shortcut is `crtl + alt + shift + v`.

##Pasting a list as a vertical vector with `vector_paste_vertical()`
Given the same types of list inputs as above, the `Paste as vector (vertical)` addin pastes the output with each element on its own line, e.g.:

```{r, eval = FALSE}
c("Mint",
  "Fedora",
  "Debian",
  "Ubuntu",
  "OpenSUSE")
```

This is much nicer for long lists. I have found this is actually the version I use more often. I recommend using `ctrl + shift + v` as keyboard shortcut.

##Pasting as a data.frame with `df_paste()`
The parser here is identical to `tribble_paste()` and has all the same type and separator guessing goodness. The difference is the output will be a formatted call to `base::data.frame()`. Some sensible line wrapping rules etc are implemented. Useful for purists and educators alike. Special thanks to Jonothan Carroll for contributing this feature.

So the Brisbane weather table from above becomes:
```{r, eval = FALSE}
data.frame(
           X = c("Partly cloudy.", "Partly cloudy.", "Possible shower.",
                 "Partly cloudy.", "Shower or two. Possible storm.",
                 "Possible shower.", "Partly cloudy.", "Mostly sunny.", "Partly cloudy.",
                 "Possible shower.", "Partly cloudy."),
    Location = c("Brisbane", "Brisbane Airport", "Beaudesert", "Chermside",
                 "Gatton", "Ipswich", "Logan Central", "Manly",
                 "Mount Gravatt", "Oxley", "Redcliffe"),
         Min = c(19, 18, 15, 17, 15, 15, 18, 20, 17, 17, 19),
         Max = c(29, 27, 30, 29, 32, 30, 29, 26, 28, 30, 27)
)
```

For a shortcut you could try `ctrl + shift + d`.
